\subsection{Quarkonia and Open Heavy Flavor}
\label{Sec:HF}
 Hadrons containing heavy quarks (charm and bottom) play a special role in hot QCD matter studies.
	The heavy-quark masses, $m_{c,b}\simeq1.5,5$\,GeV, are large compared to the temperatures typically
	realized in the medium produced in heavy-ion collisions. This has important consequences.
	First, the production of heavy quarks is essentially restricted to the initial impact of the
	incoming nuclei. After that, they ``diffuse" through the produced medium, much like Brownian
	particles. The modifications of their momentum spectra relative to the initial spectra can thus
	serve as a rather direct measure of their coupling strength to the hot medium. This is so because 
        their thermal relaxation time is increased compared to that for light particles by a factor of order 
        $m_Q/T\simeq$~5-20, implying that heavy quarks in the QGP (or hadrons in hadronic matter) are 
        not expected to fully thermalize over the course of the fireball lifetime, and thus retain a 
        ``memory" of their interaction history. Moreover, at small and intermediate momenta, the heavy-quark 
        interactions become dominantly elastic and may be amenable to a potential-type prescription. 
        In the vacuum, the potential approach is well established for the description of heavy quarkonia, 
        i.e., bound states of a heavy quark and its anti-quark. The vacuum potential is characterized 
        by a color-Coulomb interaction at short distances and a linearly rising ``string" term at 
        intermediate and large distance associated with confinement. When embedded into a QGP, the 
        properties of heavy quarkonia in the QGP will thus reflect the modifications of the in-medium
        QCD potential. With increasing temperature one expects a subsequent dissolution of the 
        bound states, as the medium-induced screening penetrates to smaller distances. However, this
        simple picture is complicated by inelastic collisions with medium particles, leading to 
        a dynamical dissociation of the bound state. These processes generate in-medium widths and
        need to be accounted for in the description of quarkonium spectral functions at finite
        temperature. In heavy ion collisions, the in-medium properties of quarkonia have to be inferred
        from their production yields and momentum spectra in the final state. A thorough interpretation
        of the data thus requires systematic analyses of the centrality, energy and species dependence
        of quarkonia. This usually requires transport approaches to calculate the time evolution of 
        quarkonium distributions including both dissociation reactions and formation processes
        through recombination of diffusing heavy quarks.   
     	In the following we briefly summarize recent progress in both open and hidden heavy flavor probes of
        hot and dense QCD matter.


\input{tex/Quarkonia}

\input{tex/OpenHeavyFlavor}
