\documentclass[12pt,letterpaper,amsmath,amssymb,final]{article}
% Fun and exciting LaTeX packages!

%Margins
\usepackage[top=1.0in, bottom=1.0in, left = 1.0in, right=1.0in]{geometry}

\usepackage{graphicx}% Include figure files 
\usepackage{tabularx}% Include figure files 
\usepackage{rotating}
\usepackage{subfig}
\usepackage{sidecap}
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bbm}% blackboard bold
\usepackage{textcomp}% I can't remember what this one does...
\usepackage{color}% Change font colors
\usepackage{gensymb}% Some nice math symbols...
\usepackage{wasysym}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{array}
\usepackage{lineno}
\usepackage{hyperref}

\setcounter{tocdepth}{3}

\setcounter{topnumber}{9}
\setcounter{bottomnumber}{9}
\setcounter{totalnumber}{9}
 
\renewcommand{\topfraction}{0.99}
\renewcommand{\bottomfraction}{0.99}
\renewcommand{\textfraction}{0.0}
\renewcommand{\floatpagefraction}{0.99}

%\linenumbers

 \def\begit{\begin{itemize}}
\def\endit{\end{itemize}}
\def\begc{\begin{center}}
\def\endc{\end{center}}
\def\begl{\begin{large}}
\def\endl{\end{large}}
\def\Begl{\begin{Large}}
\def\Endl{\end{Large}}
\def\begh{\begin{huge}\sf}
\def\endh{\end{huge}}
\def\non{\nonumber \\}
\def\nonn{\nonumber\\[0.2in]}
\def\begeq{\begin{equation}}
\def\endeq{\end{equation}}
\def\begeqar{\begin{eqnarray}}
\def\endeqar{\end{eqnarray}}  
\def\numu{\nu_\mu}
\def\nubar{{\overline\nu}}
\def\numubar{\overline\nu_\mu}
\def\nue{\nu_e}
\def\nuebar{\overline\nu_e}
\def\nutau{\nu_\tau}
\def\nutarbar{\overline\nu_\tar}
\def\lsim{\mathrel{\rlap{\lower4pt\hbox{\hskip1pt$\sim$}}
    \raise1pt\hbox{$<$}}}         %less than or approx. symbol
\def\gsim{\mathrel{\rlap{\lower4pt\hbox{\hskip1pt$\sim$}}
    \raise1pt\hbox{$>$}}}         %greater than or approx. symbol

\newcommand{\mbb}{\ensuremath{ m_{\beta\beta}}}
\newcommand{\mbbs}{\ensuremath{ m_{\beta\beta}} }
\newcommand{\nd}{$\mathrm{^{150}Nd}$}
\newcommand{\Nova}{NO\ensuremath{\nu}A}
\newcommand{\NOvA}{NO\ensuremath{\nu}A}

\begin{document}

\title{White Paper: Measuring the Neutrino Mass Hierarchy
\vspace{1.8in}}
\author{R. N. Cahn, D. A. Dwyer, S. J. Freedman$^{\dagger}$, W. C. Haxton, \vspace{0.2in}\\
R. W. Kadel, Yu. G. Kolomensky\footnote{Co-chair}, 
K. B. Luk, \vspace{0.2in}\\
P. McDonald, G. D. Orebi Gann\footnotemark[\value{footnote}], A. W. P. Poon\vspace{0.5in}\\
Lawrence Berkeley National Laboratory}

%\affiliation{Lawrence Berkeley National Laboratory}

\date{\today\vspace {1.8in}}

\maketitle
\noindent $^{\dagger}$\begin{em}{Our colleague Stuart Freedman was a member of the Neutrino Hierarchy
Working Group.  His sudden passing, on November 10, 2012, was a
terrible loss to the Nuclear Science and Physics Divisions at LBNL and
to the entire international scientific community.  It was a great personal
loss to those of us who were Stuart's friends.  Stuart set
extraordinarily high standards for scientific inquiry and integrity.
We hope that these are reflected in our report.}\end{em}
\clearpage

\begin{abstract}
\em This whitepaper is a condensation of a report by a committee
appointed jointly by the Nuclear Science and Physics Divisions at
Lawrence Berkeley National Laboratory (LBNL). The goal of this study
was to identify the most promising technique(s) for resolving the
neutrino mass hierarchy.  For the most part,  we have relied on
calculations and simulations presented by the proponents of the
various experiments.  We have included evaluations of the
opportunities and challenges for these experiments based on what is
available already in the literature. 
\end{abstract}

\setcounter{tocdepth}{2}
\tableofcontents{}

\newpage
\section{Executive Summary}
\input{ExecSV.tex}

\newpage
\section{Introduction}\label{s:intro}
\input{introSV}


\section{Related Experiments that will not Resolve the Hierarchy}\label{s:not}
\input{WontDoIt.tex}



%\newpage
%\section{Statistical Measurement of the Hierarchy}\label{s:stat}
%\input{DAD/statistics.tex}

%\newpage
%\section{Solar Neutrinos}\label{s:sol}
%\input{WH_Solar/solarnus_hierarchy}

%\newpage
%\section{Supernovae}\label{s:sn}
%\input{WH_SN/supernova}

%\newpage
%\section{Direct Neutrino-Mass Measurements}\label{s:mass}
%\input{AP/directmass}

%\newpage
%\section{Neutrinoless Double Beta Decay}\label{s:bb}
%\input{GDOG/0nubb}


\section{Long Baseline Experiments}\label{s:lb}

The long-baseline experiments have sensitivity to the neutrino mass
hierarchy, due to the interaction of neutrinos with matter as they
pass through the Earth.  The baseline itself is a critical factor in
the hierarchy sensitivity, and thus we consider the experiments in
multiple categories: near term and relatively short baseline (T2K and
\NOvA); the US-based long-baseline experiment (LBNE); and alternatives
outside the US (primarily HyperK and LBNO).  While T2K has very little
sensitivity to the hierarchy, due to the short baseline, \NOvA\ has the
potential to make a measurement at the 2--3$\sigma$ level, dependent
on the value of the CP phase parameter, $\delta_{CP}$. LBNE has
demonstrated 2$\sigma$ (3$\sigma$) sensitivity over 100\% (80\%) of
the values of $\delta_{CP}$ with 10 years of data.  LBNO has the
potential to achieve a high sensitivity on a shorter timescale, due to
the longer baseline (therefore increased sensitivity to the matter
effects).  HyperK can achieve a similar sensitivity to LBNE on a
similar timescale by combining the somewhat shorter baseline
measurement with an independent atmospheric measurement in the same
detector.  A combination of several
experiments at different baselines (e.g. T2K+\NOvA , or T2K/\NOvA +LBNE,
etc) can disentangle the competing effects of CP violation and
matter-induced neutrino-antineutrino differences, and thus improve the
constraints in the hierarchy significantly beyond any single
measurement. 

%\subsection{Can we measure the Mass Hierarchy at T2K or NOvA?}\label{s:tn}
\subsection{T2K and \NOvA }\label{s:tn}
\input{T2K_NOvA.tex}
%\newpage
\subsection{LBNE}\label{s:lbne}
\input{LBNESV}

\subsection{Non-US Options}\label{s:nonus}
\input{LBL_noUS_SV}

\section{Reactor Neutrinos}\label{s:reac}
\input{reactorSV}


\section{Atmospheric Neutrinos}\label{s:atm}
\input{atmosphericSV}


\section{Cosmology }\label{s:cosmo}
\input{cosmologySV}

%\newpage
\section{Conclusions}
\input{conclusionsSV}



\section{Acknowledgments}
We would like to thank Mark Strovink for contributions to this document, and many members of the Nuclear Science and Physics divisions at LBNL for enlightening conversations.

\newpage

\begin{thebibliography}{1}
\input{biblioSV}
\end{thebibliography}


\end{document}

























